---
title: "Analysis"
author: "Charles Alba"
date: "2022-11-05"
output:
  pdf_document: default
  html_document: default
---

# Preparing the data. 


```{r}
### load library
library(splm)
library(Ecdat)
library(data.table)
library(sf)
library(spdep)
library(rgdal)
library(naniar)
library(stringr)
library(dplyr)
library(tidyr)
library(bit64)
library(stargazer)
library(plm)
library(sjPlot)
library(ggplot2)
library(boot)
library(car)


# Reading dataset 
data<-read.csv("data.csv")

# cleaning the dataset 
data$date<-(as.Date((as.character(data$date_range_start)), format='%Y-%m-%d'))
data$county<-str_pad(as.character(data$county), 5, pad = "0")


```


# Calculating unhealthy food reliance & preparing our data for analysis 

```{r}

# Calculating unhealthy food reliance

## Calculating convenience-store reliance
data$cs<-data$convenient_store_count/((data$convenient_store_count)+(data$grocery_count))

## Calcularing fast-food reliance 
data$ff<-(data$fast_food_count) / (data$fast_food_count+data$full_service_count)

## calculating covid incidence rate. 
data$cv<-((data$monthly_cases)/(data$Total_Population))

# removing irrelevant columns
data<-subset(data,select=-c(Total_Population,date_range_start,convenient_store_count,grocery_count,fast_food_count,full_service_count,monthly_cases,Total_Population,monthly_deaths,Geo_QName))
data<-data.table(data)


# ensuring that in the logit() function 0's and 1's are accepted
data[is.na(data)] = 0
data$cs[data$cs==0]<-0.00001
data$cs[data$cs==1]<-0.99999
data$ff[data$ff==0]<-0.00001
data$ff[data$ff==1]<-0.99999
data<-data.frame(data)

  
```

# Establishing the model. 

```{r}

# This function performs our models. 
fixed_model<-function(formula,d=data){
  fixed<- plm(formula = formula, data = d, index = c("county","date"), model = "within",effect = "time")
  (summary(fixed)) 
}

# This function generates fitted-values and can be used for our interaction plot 
predictions<-function(formula,d){
  d<-group_by(d) %>%mutate(check = ntile(cv, 100))
  
  d$covid <-ifelse(d$check >= 90, "High (>90th percentile)", # classifying COVID into high, moderate and no/few incidence rates based on quantiles, to be used later for our interaction plots. 
    ifelse(d$check > 5 & d$check < 90, "Moderate COVID cases",
      ifelse(d$check > 0 & d$check <= 5, "Low (<5th percentile)", NA)))
  d$predicts<-predict(plm(formula = formula, data = d, index = c("county","date"), model = "within",effect = "time"))
  d<-d[!(d$covid=="Moderate COVID cases"),]
  return(d)
  #return(data)
}
```

# performing the analysis

## Performing the analysis for convenient-store reliance. 

```{r}
# performing the analysis here. 
fm_w <- logit(cs)~ cv + percent_black + percent_hispanic + percent_native + percent_asian + percent_black:cv + percent_hispanic:cv + percent_asian:cv + percent_native:cv
fixed_model(fm_w)
```



### Making Interaction plots based on our results above. 

#### Interaction plots for African-Americans

```{r}
data_conv<-predictions(fm_w,data)
ggplot(data=data_conv, aes(x=percent_black*100, y=inv.logit(predicts)*100, group=covid))+geom_smooth(aes(color=covid), method="gam", level=0.95)+ylab("convenience-store reliance (%)")+xlab("% of African Americans")+ scale_color_discrete(name = "COVID incidence rate")+geom_point(aes(color=covid),alpha=0.01)+ theme_linedraw()+ylim(0,50)
```

#### Interaction plots for Asians. 

```{r}
ggplot(data=data_conv, aes(x=percent_asian*100, y=inv.logit(predicts)*100, group=covid))+geom_smooth(aes(color=covid), method="gam",level=0.95)+ylab("convenience-store reliance (%)")+xlab("% of Asian Americans")+geom_point(aes(color=covid),alpha=0.02)+ scale_color_discrete(name = "COVID incidence rate")+ theme_linedraw()+ylim(0,50)
```

#### Interaction plots for Native Americans. 


```{r}
ggplot(data=data_conv, aes(x=percent_native*100, y=inv.logit(predicts)*100, group=covid))+geom_smooth(aes(color=covid), method="gam",level=0.95)+ylab("convenience-store reliance (%)")+xlab("% of Native Americans")+geom_point(aes(color=covid),alpha=0.02)+ scale_color_discrete(name = "COVID incidence rate")+ theme_linedraw()+ylim(0,50)
```


#### Combining graphs for aesthetic purposes. 

```{r}

# elongating the data to make it easier to plot the different races in the same plot with ggplot
elongate_data<-function(data_conv){
  data_long<-subset(data_conv,select=c(predicts,percent_asian,percent_native,percent_black,percent_hispanic,county,date,covid))
  data_long <- melt(data_long, id = c("predicts","date","county","covid"))
  data_long$variable <- gsub('percent_black', 'African american', data_long$variable)
  data_long$variable <- gsub('percent_asian', 'Asian american', data_long$variable)
  data_long$variable <- gsub('percent_native', 'Native american', data_long$variable)
  data_long$variable <- gsub('percent_hispanic', 'Hispanic', data_long$variable)
  colnames(data_long)[colnames(data_long) == "variable"] <- "Race" # Rename column
  data_long$covid<-factor(data_long$covid, levels=c("Low (<5th percentile)","High (>90th percentile)"))
  return(data_long)
}

data_long<-elongate_data(data_conv)


ggplot()+geom_point(data=data_long,aes(x=value*100, y=inv.logit(predicts)*100,shape=covid,color=Race),alpha=0.02,legend=FALSE)+
  geom_point(data=data_long, aes(x = value - 999999, y=inv.logit(predicts)*100  - 999999, shape=covid,colour = Race), alpha=1.0)+geom_smooth(data=data_long,aes(x=value*100, y=inv.logit(predicts)*100,color=Race,lty=covid), method="gam",level=0.95)+ylab("Convenience store reliance (%)")+xlab("% of race")+ scale_linetype_discrete(name ="COVID incidence rate")+scale_color_discrete(name="County-level race percentage")+ theme_linedraw()+ylim(0,55)+xlim(0,100)

ggplot(data=data_long, aes(x=value*100, y=inv.logit(predicts)*100, col=Race,lty=covid))+geom_smooth(aes(color=Race,lty=covid), method="gam",level=0.95)+ylab("Convenience store reliance (%)")+xlab("% of race")+ scale_linetype_discrete(name = "COVID incidence rate")+ theme_linedraw()+scale_linetype_manual(values=c("solid", "twodash"))+ggtitle("Figure 2a.\nThe effect of racial demographics and COVID-19\non convenience store reliance")

```




## Performing analysis for Fast-food reliance

```{r}
fm_w <- logit(ff)~ cv + percent_black + percent_hispanic + percent_native + percent_asian + percent_black:cv + percent_hispanic:cv + percent_asian:cv + percent_native:cv
fixed_model(fm_w)
data_fast_food<-predictions(fm_w,data)
```

### Making interaction plots based on the results above

#### Interact plots for African-Americans 


```{r}
ggplot(data=data_fast_food, aes(x=percent_black*100, y=inv.logit(predicts)*100, group=covid))+geom_smooth(aes(color=covid), method="gam",level=0.95)+ylab("convenience-store reliance (%)")+xlab("% of African Americans")+ scale_fill_discrete(name = "COVID incidence rate")+geom_point(aes(color=covid),alpha=0.02)+ scale_color_discrete(name = "COVID incidence rate")+ theme_linedraw()
```


#### Interact plots for Asian-Americans 

```{r}
ggplot(data=data_fast_food, aes(x=percent_asian*100, y=inv.logit(predicts)*100, group=covid))+geom_smooth( aes(color=covid), method="gam",level=0.95)+ylab("convenience-store reliance (%)")+xlab("% of Asian Americans")+geom_point(aes(color=covid),alpha=0.02)+ scale_color_discrete(name = "COVID incidence rate")+ theme_linedraw()
```

#### Interact plots for Native Americans. 

```{r}
ggplot(data=data_fast_food, aes(x=percent_native*100, y=inv.logit(predicts)*100, group=covid))+geom_smooth(aes(color=covid), method="gam")+ylab("convenience-store reliance (%)")+xlab("% of Native Americans")+ scale_color_discrete(name = "COVID incidence rate")+geom_point(aes(color=covid),alpha=0.02)+ theme_linedraw()
```


#### Combining Graphs for aesthetic purposes. 

```{r}
data_long<-elongate_data(data_fast_food)


ggplot()+geom_point(data=data_long,aes(x=value*100, y=inv.logit(predicts)*100,shape=covid,color=Race),alpha=0.015,legend=FALSE)+
  geom_point(data=data_long, aes(x = value - 999999, y=inv.logit(predicts)*100  - 999999, shape=covid,colour = Race), alpha=1.0)+geom_smooth(data=data_long,aes(x=value*100, y=inv.logit(predicts)*100,color=Race,lty=covid), method="gam",level=0.95)+ylab("fast-food reliance (%)")+xlab("% of Race")+ scale_linetype_discrete(name ="COVID incidence rate")+ theme_linedraw()+ylim(40,90)+xlim(0,100)

ggplot(data=data_long, aes(x=value*100, y=inv.logit(predicts)*100, col=Race,lty=covid))+geom_smooth(aes(color=Race,lty=covid), method="gam",level=0.95)+ylab("Fast food reliance (%)")+xlab("% of race")+ scale_linetype_discrete(name = "COVID incidence rate")+ theme_linedraw()+scale_linetype_manual(values=c("solid", "twodash"))+ggtitle("Figure 2b.\nThe effect of racial demographics and COVID-19\non fast-food reliance")

```






